### 2018-06-09 MSc_Marko_Docevski_tekst.tex
Updates from ShareLaTeX

\cite{Bretan2016}
The work presented in \cite{Bretan2016} approaches music generation as a word-level language model, where each unit (composed of 1 to 4 bars of music) is treated as word. The words are embedded in a descriptor vector. A library of word (or units) is then created from the complete set of embeddings. A LSTM network learns sequences of the embeddings. Music is generated by first sequencing embeddings, which are then decoded to actual note sequences by selecting the most appropriate representative out of the library of units. We used character level embeddings in out second experiment.

The process for identifying good candidates is based on the assumption that two contiguous units,(un−1,un), should share characteristics in a higher level musical semantic space (semantic relevance) and the transition between the last and first notes of the first and second units respectively should be likely to occuraccording to a model (concatenation). This general idea is visually portrayed in Figure 4. We use a DSSM based on BOW-like features to model the semantic relevance between two contiguous units and a note-level LSTM to learn likely note sequences (where a note contains pitch and rhythm information). For training these models we use the same dataset described in the previous section. However, in order to ensure that the model learns sequences and relationships that are musically appropriate we can only augment the dataset by transposing the pieces to different keys. Transposing does not compromise the original structure, pitch intervals, or rhythmic information within the data, however, the other transformations do affect these musical attributes and such transformations should not be applied for learning the parameters of these sequential models. However, it is possible to use the original unit library (including augmentations) when selecting units during generation.
The process for identifying good candidates is based on the assumption that two contiguous units, (un−1,un), should share characteristics in a higher level musical semantic space (semantic relevance) and the transition between the last and first notes of the first and second units respectively should be likely to occur according to a model (concatenation). This general idea is visually portrayed in Figure 4. We use a DSSM based on BOW-like features to model the semantic relevance  between two contiguous units and a note-level LSTM to learn likely note sequences (where a note contains pitch and rhythm information).
We use a multi-layer LSTM to learn a note-to-note level model (akin to a character level language model). The aim of the concatenation cost is to compute a score evaluating the transition between the last note of the unit, un−1,xT , and the first note of the unit, un,yT . By using an LSTM it is possible to include additional context and note dependencies that exist further in the past than The sequence length, T = 36, was chosen because it is roughly the average number of notes in four measures of music (from our dataset). Unlike the DSSM, which computes distances based on information from a fixed number of measures, the context provided to the LSTM is fixed in the number of notes. This means it may look more or less than four measures into the past. In the scenario in which there is less that 36 notes of available context the sequence is zero padded.
A ranking process that combines the semantic relevance and concatenation cost is used to perform unit selection. The ranking process is performed in four steps:
1. Rank all units according to their semantic relevance with an input seed using the feature space learned by the DSSM.
2. Take the units whose semantic relevance ranks them in the top 5percent and re-rank based on their con- catenation cost with the input.
3. Re-rank the same top 5percent based on their combined semantic relevance and concatenation ranks.
The model’s ability to choose good units can be evaluated using a ranking test. The task for the model is to predict the next unit given a never before seen four measures of music (from the held out test set). The prediction is made by ranking 50 candidates in which one is the truth and the other 49 are units randomly selected from the database. We repeat the experiments for musical units of different lengths including four, two, and one measures. The results are reported in the table below and they are based on the concatenation cost alone (LSTM), semantic relevance (DSSM), and the combined concatenation and semantic relevance using the selection process described above (DSSM+LSTM).
Subjective Evaluation
A subjective listening test was performed. Participants included 32 music experts in which a music expert is defined as an individual that has or is pursuing a higher level degree in music, a professional musician, or a music educator. Four systems were evaluated. Three of the systems employed unit selection using units of four, two, and one measures. The fourth system used the note-level LSTM to generate each note at a time. Each participant was asked to evaluate 10 samples that were randomly selected from the original 60, thus, all participants listened to music generated by the same four systems, but the actual musical content and order randomly differed from participant to participant. The tests were completed online with an average duration of roughly 80 minutes

@@

\cite{Boulanger-Lewandowski2012} Polyphonic music is a sequence od high-dimensional objects. Simply predicting the expected value at the next time step given an observed value of a previous timestep is not always satisfactory. The conditional distribution at each time step is often multi-modal, so prediction of the conditinal distribution of the next time step given previous time steps can be prefered. The occurence of a particular note at a time step obviously influences the probability of other notes occuring at the same time. In other words notes appear together in corelated patterns, or simulateities, that cannot be conveniently described by a typical RNN architecture designed for multiclass classificaiton, because enumeration of all possible configurations of the variable to predict can be very expensive. This lead the researchers towards energy based models, which allow expression of the negative log-likelihood of a given configuration by an arbitrary energy function. RBM is a prime candidate for this. Temporal RBM - using the RBM to represent complex distributions at each timestep, with parameters that depend on the ones fom the previous step. Combining the desirable characteristics of RNNs and RBMs is not trivial.. The RTRBM (Recurrent Temporal RBM) is a model that allows for exact inference and effective training with contrastive divergence (CD). RNN-RBM is a generalization of RTRBM put forward in this paper, should provide more freedom to describe the temporal dependencies. Musical models mostly focus on the basic components of western music, harmony and rhythm, and are trained to predict the pattern of notes (simultaneities) to be played together in the next time interval, given the previous ones. Two elements characterize the qualitative performance of a model: temporal dependencies and chord conditional distributions. They model unconstrained polyphonic music in the piano-roll representation, i.e. as a binary matrix specifying precisely which notes occur at each time step. The objective of polyphonic transcription is to determine the underlying notes of a polyphonic audio signal without access to its score. The neural autoregressive distribution estimator (NADE) is a tractable model inspired by the RBM and specializing (with tying constraints) an earlier model for the joint distribution of high-dimensional variables. The RTRBM is a sequence of conditional RBMs (one at each time step) whose parameters bv(t),bh(t) ,W(t) are time-dependent and depend on the sequence history at time t, denoted where is the mean-field value of h(t). The RTRBM can be understood as a sequence of conditional RBMs whose parameters are the output of a deterministic RNN, with the constraint that the hidden units must describe the conditional distributions and convey temporal information. This constraint can be lifted by combining a full RNN with distinct hidden units h(t) with the RTRBM graphical model as shown in Figure 2(b). We call this model the RNN-RBM. The RNN-NADE, obtained by substituting NADEs for RBMs, allows for exact gradient computation. Pre training proved to improve the overall network performance of both RTRBM and RNN-RBM. Hessian Free optimization significantly impacts the density estimation and prediction perfromance of RNNs. Learning rates and other hyperparameters can and maybe should be specified differently for the RNN and RBM parts. They can have a big impact on prediciton perfromance. This stresses the importance of pretraining and finetuning of the RNN, and the aditional advantage of using HF. Although frame lavel NADEs are slightly less powerful than RBMs, their desirable properties make them perform more robustly in the combined RNN-NADE morel. While note correlations are obviously neglected in the simpler models , RBM-based models learned basic harmony rules, melody lines and local temporal coherence. However the model does not capture long term dependencies.

\cite{Boulanger-Lewandowski2014} Boulanger-Lewandowski et al. in \cite{Boulanger-Lewandowski2014} attempted to combine the RNNs' ability to model sequences and Restricted Boltzmann Machines' (RBMs) ability to model probability distributions and apply them to music generation, in an architecture dubbed RNN-RBM. In this architecture the RBM part models note co-occurrence probability distributions at each time-step and the RNN part models the temporal sequences. We aim to achieve polyphony by modeling  the probability of each output of the network independently of one another, even though realistically they are not, and try to avoid using a hybrid solution like RNN-RBN, due to it being difficult to train and computationally expensive.

>>>>

Врз основа на техники за обработка на природни јазици може да се направат и модели на ниво на збор. Ваков пристап си има свои ограничувања бидејќи се зголему драстично бројот на основни елементи, има многу повеќе зборови одошто букви и симобили во јазициите, и нивните фрекфенции на појавување опаѓаат драстично. За да се надмине ова се користат различни техники за намалување на бројот на варијабли во системот, т.е. намалување на бројот на зборови, групирање на зборовите според функција, споделување на веројантности помеѓу зборовите и трансформации во помалку димензионален простор. Токму ваков пристап може да се види во \cite{Bretan2016}, каде што авторите третираат исечоци од 1 до 4 такта како зборови. Бидејќи има огромен број на можни варијации на таквите исечоци истите ги смалуваат во помалку димензионален простор користејќи ја техниката на вградување во латентен простор (анг. latent space embedding). Вградувањето го вршат врз основа на сет од дескриптори со кои ги опишуваат исечоците. Со ова се тобива структура налик на хеш табела, само со повеќе можни резултати при процесот на декодирање од елемент во табела во исечок. Тренираат два модела, еден што учи секвенци од репрезентации на зборовите, а друг што ги учи песните на ниво на буква и е наменет за решавање на проблемот на декодирање, т.е. ги гледа работиве помеѓу тековната секвенца и сите потенцијални декодирани исечоци, и помага во изворот на најсоодветниот кандидат. Намената на моделот на ниво на збор е да учи музички зависности на подолг временски период, во обид да се научи структура на цела песна, додека другиот модел ги учи локалните зависности помеѓу нотите. На крај извршиле субјективна евалуација од страна на 32 музички експеримент, коишто ги рангирале резултатите од повеќе инстанци на моделот со различн параметри.

@@

Користењето на јазични модели на ниво на бука овозможува едноставно моделирање на монофона музика, или музика која во секој момент има највеќе една активна нова. При моделирање на акорди со пиано лента се појавува проблем на енумерирање сите можни конфигурации на ноти што може да се појават, пр. за гитара има грубо $2*24^6$ можни конфигурации на ноти. Еден пристап за намалување на проблемите што доаѓаат од зголемена димензионалност може да се види во \cite{Boulanger-Lewandowski2012, Boulanger-Lewandowski2014, Goel2014}, каде што авторите користат хибридна архитектура, изграден од рекурентни невронски слоеви во различни комбинации со енергетски модели. Идејата е да се искористат особеностите на енергетските модели, како што се RBM (Restricted Boltzmann Machines - Ограничени Болзманови Машини) и NADE (Neural Autoregressive Distribution Estimator), да моделираат веројатностни дистрибуции, кон моделирање на истовремено појавување на нотите во акорди. Од \cite{Boulanger-Lewandowski2012, Boulanger-Lewandowski2014} произлегуваат моделите: RTRBM (Recurrent Temporal RBM - Рекурентна Темпорална Ограничена Болзманова Машина), RNN-RBM, RNN-NADE, од кои RNN-NADE се покажал како наједноставен за тренирање и со најдобри генеративни резултати и најекспресивен, иако од енергетските модели NADE има релативно помала експресивна моќ. Моделот може да се разбере како низа од условени енергетски слоеви, по еден за секој временски чекор што го обработува мрежата, со излез во детерминистички рекурентен невронски слој. Наголем проблем на овој тип на модели се покажала осетливоста на иницијална состојба, што многу ја потенцира важноста на пред-тренирање при нивна употреба. Моделот се покажал успешен во моделирање на локални зависности, хармониски правила и кратки мелодиски линии, меѓутоа не може да моделира долговременски зависности.
